Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-04-18 VegOzone.tex
Rewrite the last paragraph of section 4

A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics. This is because they are able to account particular species characteristics (i.e. physiology and phenology) as well as environmental conditions that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}. These flux based metrics have been shown to be better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2010,ICP:MappingManual2017}. Most of these previous studie have focussed on the Mediterranean where soil moisture deficits are thought to be most influential in decoupling concentration from uptake. Relatively few studies have explored the situation in Northern European climates where low temperatures may be considered more likely to influence ozone uptake \citep[e.g.]{iF:Juran2018}.

@@

An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover biomonitor. This study clearly demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of vegetation to ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe. A similar biomonitoring study was conducted at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether ozone was capable of inducing damage to vegetation growing under northern Fennoscandia conditions. In 2018, we indeed observed ozone damage on clover and tobacco. In contrast, no such damage was found on the clovers in 2019.

@@

The main species or PFTs, we deduced on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}, are perennial grassland, downy birch, and Norway spruce. Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2010,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following. A comprehensive list of model parameters can be found in Supplement~Table~S1.
Initial test simulations with the MM PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}). Only during an extended warm period in late July did we find substantial stomatal conductance. In ecological terms this would mean that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality. We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance. Therefore, we propose bespoke parameterizations which may describe an actual acclimation of these species to the climatic conditions in the target region.

@@

The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation. In summary, we identify the following main effects on the model predicted ozone uptake with respect to the MM default parameterization:
\begin{itemize}
  \itemsep0pt
\item Increasing cold tolerance/heat intolerance $\Rightarrow$ increased uptake regardless of individual year,
\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
\item an earlier assumed start of GS $\Rightarrow$ increased uptake,
\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
\item lower/higher light threshold $\Rightarrow$ increased/reduced uptake,
\item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
\item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought effects.
\end{itemize}
The exact order of magnitude of these effects varies for individual species or PFTs and between years, but the predicted ozone uptake for the bespoke parameterization is in any case larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here. Although our temperature acclimations are hypothetical, this indicates a possibly considerable underestimation of ozone risk for subarctic species. Especially, an earlier GS due to climate change, that is overlapping with the ozone spring peak, has the potential to increase the ozone damage risk for vegetation with an acclimation to cold climates. Drought effects (SWP) do only matter if a bespoke GS is not taken into account. We will look at the specific species or PFTs in more detail in the following.

For downy birch (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Downy birch displays the largest variance of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the sensitivity to the light threshold is especially pronounced, e.g. $\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of the boreal type without bespoke GS in 2019. Without bespoke GS, 2018 displays a higher ozone damage risk than 2019. However, taking the later start and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type. This can be partly explained by very high light response factors (over $f_\mathrm{light}>0.8$) in the beginning of the GS in these simulations. Whether this is a bug or a feature of the $\mathrm{DO_3SE}$ model is not yet clear. SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Norway spruce is the least sensitive to an acclimation in temperature and therefore the least biased. Norway spruce displays a medium sensitivity to the light threshold, e.g. $\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018. Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL. Perennial grassland shows the smallest ozone uptake, but a similarly large bias as downy birch. We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the boreal type. Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

@@

The estimated biomass reduction in \unit{\%} for the bespoke temperature response function of type boreal in accordance to \citet{ICP:MappingManual2017, ESPR:Hayes2021} is shown in Table~\ref{tab:biomass_reduction}. The estimated uncertainty range corresponds to the deviation given by the $\pm 20\,\unit{\%}$ variations of the light threshold and SWP=on. For downy birch and perennial grassland, substantial reductions in total biomass are predicted for any of the two years. Norway spruce is least affected. The systematic uncertainties for downy birch are one order of magnitude larger than for Norway spruce and perennial grassland. For all species or PFTs the interannual variability is larger than the systematic uncertainties estimated for a specific acclimation and is of the same order of magnitude as the divergence between the hypothetical acclimations. For comparison, we averaged the corresponding biomass reduction for the MM temperature and light response functions with both SWP on/off but with bespoke GS over both years and find a non negligible, which means above systematic uncertainties, low-bias for both downy birch and perennial grassland with respect to the MM default parameterization. For Norway spruce, the default MM parameterization average lies within the given systematic uncertainty range, this may reflect the more comprehensive research on this species or PFT.

\begin{table}[t]
  \caption{Estimated total biomass reduction in \unit{\%} for bespoke temperature response function of type boreal with bespoke GS and SWP=off. The uncertainty ranges reported here correspond to maximum a divergence by varying the light threshold and turning SWP sensitivity on in $\mathrm{DO_3SE}$. For comparison, the corresponding biomass reduction for the MM temperature and light response functions, SWP on/off, and bespoke GS averaged over both years is shown with its standard deviation.}
  \label{tab:biomass_reduction}
\begin{tabular}{lllll}
\tophline
Year & \multicolumn{4}{c}{PFT}\\
& Downy birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
\middlehline
2018 & $15.5^{+(1.9...2.1)}_{-(0.8...1.5)}$ & $2.5^{+0.2}_{-0.2}$ & $9.7^{+0.2}_{-0.2}$ & $^*12.4^{+0.2}_{-0.2}$\\
2019 & $17.4^{+2.5}_{-1.8}$ & $3.0^{+0.2}_{-0.3}$ & $10.5^{+(0.1...0.2)}_{-(0.0...0.1)}$ & $^*13.7^{+(0.1...0.3)}_{-(0.1...0.3)}$\\
\middlehline
$\left<\mathrm{MM}\right>$ & $11.2\pm 1.1$ & $2.31\pm 0.04$ & $7.5\pm 0.9$ & $^*9.6\pm 1.4$\\
\bottomhline
\end{tabular}
\belowtable{$^*$ above ground biomass} % Table Footnotes
\end{table}

@@

To estimate potential biases in assessment of ozone induced risk on Fennoscandia vegetation, we defined hypothetical temperature response functions representing different levels of assumed acclimation to cold climates. We assume northern Fennoscandia vegetation to be more tolerant towards cold temperatures and less tolerant to heat. Based on probability density functions (PDF) of observed light and temperature, we assume that an acclimation ought to increase the enclosed area between the respective response functions and the PDF. These hypothetical species or PFTs still need verification by experimental data. Our results, however, allowed for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric. In particular, we identified the following main effects on the model predicted ozone uptake with respect to the default parameterization:
\begin{itemize}
  \itemsep0pt
\item Increasing cold tolerance/heat intolerance $\Rightarrow$ increased uptake regardless of individual year,
\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
\item an earlier assumed start of growing season $\Rightarrow$ increased uptake,
\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
\item lower/higher light threshold $\Rightarrow$ increased/reduced uptake,
\item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
\item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought.
\end{itemize}
The exact order of magnitude of these effects varies for individual species or PFTs and between years. Only accounting for acclimation, the predicted ozone uptake is typically larger than for default parameterizations. This indicates a considerable underestimation of ozone risk for subarctic species, especially for downy birch and perennial grassland. However, taking bespoke length and timing of growing season into account the opposite is be the case under certain conditions, e.g. a later start of the growing season with respect to the ozone spring peak for downy birch and perennial grassland. Simulations of Norway spruce with bespoke (shorter than default) growing season show only little sensitivity to acclimations and inter-annual variation of meteorological conditions. We have not assessed the effect of VPD response function here, but our climatological data show that conditions at Svanhovd are shifting towards a more VPD limited regime.

>>>>

A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics. This is because they are able to account particular species characteristics (i.e. physiology and phenology) as well as environmental conditions that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}. These flux based metrics have been shown to be better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017}. Most of these previous studie have focussed on the Mediterranean where soil moisture deficits are thought to be most influential in decoupling concentration from uptake. Relatively few studies have explored the situation in Northern European climates where low temperatures may be considered more likely to influence ozone uptake \citep[e.g.]{iF:Juran2018}.

@@

An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover biomonitor. This study clearly demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe. A similar biomonitoring study was conducted at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether ozone was capable of inducing damage to vegetation growing under northern Fennoscandia conditions. In 2018, we indeed observed ozone damage on clover and tobacco. In contrast, no such damage was found on the clovers in 2019.

@@

The main species or PFTs, we deduced on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}, are perennial grassland, downy birch, and Norway spruce. Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following. A comprehensive list of model parameters can be found in Supplement~Table~S1.
Initial test simulations with the MM PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}). Only during an extended warm period in late July did we find substantial stomatal conductance. In ecological terms this would mean that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality. We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance. Therefore, we propose bespoke parameterizations which may describe an actual acclimation of these species to the climatic conditions in the target region.

@@

The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation. In summary, we identify the following main effects of the bespoke parameterizations of the $DO_3SE$ model with respect to the MM default parameterization on ozone uptake:
\begin{itemize}
  \itemsep0pt
\item Increasing cold tolerance/heat intolerance $\Rightarrow$ increased uptake in both probed years,
\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the increase reported above,
\item an earlier assumed start of GS $\Rightarrow$ increased uptake,
\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
\item lower/higher extent of stomatal opening at low light intensities $\Rightarrow$ increased/reduced uptake,
\item symmetric variation of the extent of stomatal opening at low light intensities $\Rightarrow$ asymmetric response in POD,
\item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought effects.
\end{itemize}
The exact order of magnitude of these effects varies for individual species or PFTs and between years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here. Although our temperature acclimations are hypothetical, they suggest considerable underestimations of ozone risk for subarctic species when compared to generic MM parameterisations. In particular, an earlier GS due to climate change, that is overlapping with the ozone spring peak, has the potential to increase the risk of ozone damage for vegetation with an acclimation to cold climates. Drought effects (SWP) do only matter if a bespoke GS is not taken into account. We will look at the specific species or PFTs in more detail in the following.

For downy birch (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Downy birch displays the largest variance of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the sensitivity to the light threshold is especially pronounced, e.g. $\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of the boreal type without bespoke GS in 2019. Without bespoke GS, 2018 displays a higher ozone damage risk than 2019. However, taking the later start and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type. This can be partly explained by very high light response factors (over $f_\mathrm{light}>0.8$) in the beginning of the GS in these simulations. Whether this is a bug or a feature of the $\mathrm{DO_3SE}$ model is not yet clear. SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS. The temperature acclimation dependent bias in ozone uptake deduced for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Norway spruce is the least sensitive to an acclimation in temperature and therefore the least biased compared to the generic default MM parameterization. Norway spruce displays a medium sensitivity to the variation in the light threshold, e.g. $\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018. Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL. Perennial grassland shows the smallest ozone uptake, but a similarly large bias as downy birch. We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the boreal type. Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

@@

We estimated the biomass reductions in accordance to the CLs in \citet{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature acclimation (\emph{boreal}) for each bespoke PFT. The estimates are listed in Table~\ref{tab:biomass_reduction}. The reported range in uncertainty corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of the extent of stomatal opening at low light intensities with SWP=on. For downy birch and perennial grassland, substantial reductions in total biomass in comparison with the default generic MM parameterizations are predicted for both years. Norway spruce is the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties. This may reflect the more comprehensive research on this species or PFT. The systematic deviation for downy birch is one order of magnitude larger than for Norway spruce and perennial grassland.

From the difference between 2018 and 2019, we infer a substantial interannual variability. The interannual variability is estimated to be larger than our reported systematic deviations and can be of the same order of magnitude as the effect of the temperature acclimation. This means that the choice of temperature acclimation has the potential to either revoke or amplify the risk of ozone damage for vegetation. However, the risk of ozone damage in downy birch and perennial grassland in comparison with the generic default MM parametrizations is increased, regardlessly.

\begin{table}[t]
  \caption{Estimated total biomass reduction in \unit{\%} for temperature acclimation \emph{boreal} with bespoke GS and SWP=off. The uncertainty ranges reported here correspond to maximum a divergence deducted from varying the extent of stomatal opening at low light intensities and SWP=on. For comparison, the corresponding biomass reduction for the generic default MM averaged over both years is shown. The reported standard deviation is computed from all sensitivity simulations ($\pm 20\,\unit{\%}$ in extent of stomatal opening at low light, SWP on/off, and bespoke GS).}
  \label{tab:biomass_reduction}
\begin{tabular}{lllll}
\tophline
Year & \multicolumn{4}{c}{PFT}\\
& Downy birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
\middlehline
2018 & $15.5^{+(1.9...2.1)}_{-(0.8...1.5)}$ & $2.5^{+0.2}_{-0.2}$ & $9.7^{+0.2}_{-0.2}$ & $^*12.4^{+0.2}_{-0.2}$\\
2019 & $17.4^{+2.5}_{-1.8}$ & $3.0^{+0.2}_{-0.3}$ & $10.5^{+(0.1...0.2)}_{-(0.0...0.1)}$ & $^*13.7^{+(0.1...0.3)}_{-(0.1...0.3)}$\\
\middlehline
$\left<\mathrm{MM}\right>$ & $11.2\pm 1.1$ & $2.31\pm 0.04$ & $7.5\pm 0.9$ & $^*9.6\pm 1.4$\\
\bottomhline
\end{tabular}
\belowtable{$^*$ above ground biomass} % Table Footnotes
\end{table}

@@

To estimate potential biases in assessment of ozone induced risk on Fennoscandia vegetation, we defined hypothetical temperature response functions representing different levels of assumed acclimation to cold climates. We assume northern Fennoscandia vegetation to be more tolerant towards cold temperatures and less tolerant to heat. Based on probability density functions (PDF) of observed light and temperature, we assume that an acclimation ought to increase the enclosed area between the respective response functions and the PDF. These hypothetical species or PFTs still need verification by experimental data. Our results, however, allowed for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric. In particular, we identified the following main effects on the model predicted ozone uptake with respect to the default parameterization:
\begin{itemize}
  \itemsep0pt
\item Increasing cold tolerance/heat intolerance $\Rightarrow$ increased uptake regardless of individual year,
\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
\item an earlier assumed start of growing season $\Rightarrow$ increased uptake,
\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
\item lower/higher extent of stomatal opening at low light intensities $\Rightarrow$ increased/reduced uptake,
\item symmetric variation of the extent of stomatal opening at low light intensities $\Rightarrow$ asymmetric response in POD,
\item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought.
\end{itemize}
The exact order of magnitude of these effects varies for individual species or PFTs and between years. Only accounting for acclimation, the predicted ozone uptake is typically larger than for default parameterizations. This indicates a considerable underestimation of ozone risk for subarctic species, especially for downy birch and perennial grassland. However, taking bespoke length and timing of growing season into account the opposite is be the case under certain conditions, e.g. a later start of the growing season with respect to the ozone spring peak for downy birch and perennial grassland. Simulations of Norway spruce with bespoke (shorter than default) growing season show only little sensitivity to acclimations and inter-annual variation of meteorological conditions. We have not assessed the effect of VPD response function here, but our climatological data show that conditions at Svanhovd are shifting towards a more VPD limited regime.
